---
title: "Cumulative"
format: html
author: "Shuntaro Kawakami"
execute: 
  echo: false
---


```{r}
library(MASS)
library(tidyverse)
library(primer.data)
library(broom) 
library(easystats)
library(marginaleffects)
x <- ces |>
  filter(year == 2020) |>
  select(approval, ideology, education) |>
  drop_na() |> 
  filter(! ideology %in% "Not Sure") |> 
  mutate(ideology = fct_drop(ideology))
x |> ggplot(aes(x = ideology, fill=approval)) + geom_bar() + labs(title = "Presidential Approval by Ideology", subtitle = "Different ideology groups show different approval levels", x="Ideology", y="Probability")
fit_approval <- polr(approval ~ ideology + education,
                      data = x)
fit_approval |> tidy(conf.int = TRUE)
tidy_approval <- read_rds("data/tidy_approval.rds")

```
We are trying to find out approval rating of president Trump in 2020 based on ideology of individual.  Using data from the 2020 Cooperative Election Study, one of the largest political surveys in the U.S., we explore how presidential approval varies across different ideological groups.

Public opinion about political leaders often depends on where people fall on the ideological spectrum. Using data from the 2020 Cooperative Election Study, one of the largest political surveys in the U.S., we explore how presidential approval varies across different ideological groups. One concern is that survey respondents might be systematically different from other Americans.

\[
\log\left( \frac{P(Y \leq k)}{P(Y > k)} \right)
= \theta_k - (\beta_1 \cdot \text{ideology} + \beta_2 \cdot \text{education})
\]

$$P(Y \leq k) = \frac{1}{1 + e^{-(\alpha_k - \beta_1 \cdot \text{ideology} - \beta_2 \cdot \text{education})}}$$

with $Y \sim \text{Ordinal}(\boldsymbol{\rho})$ where $\boldsymbol{\rho} = (\rho_1, \rho_2, \ldots, \rho_K)$ are derived from the cumulative probabilities above.

$$
\hat{Y} = 1.25 + 0.48 \cdot \text{ideology} - 0.32 \cdot \text{education} + 0.15 \cdot \text{income}
$$

$$\begin{aligned}
\log\left[ \frac { P( \operatorname{approval}  \leq  \operatorname{Strongly\ Disapprove} ) }{ 1 - P( \operatorname{approval}  \leq  \operatorname{Strongly\ Disapprove} ) } \right] &= 2.26 + 0.61(\operatorname{ideology}_{\operatorname{Liberal}}) + 2.4(\operatorname{ideology}_{\operatorname{Moderate}}) + 4.46(\operatorname{ideology}_{\operatorname{Conservative}}) + 5.45(\operatorname{ideology}_{\operatorname{Very\ Conservative}}) - 0.1(\operatorname{education}_{\operatorname{High\ School\ Graduate}}) - 0.31(\operatorname{education}_{\operatorname{Some\ College}}) - 0.32(\operatorname{education}_{\operatorname{2-Year}}) - 0.52(\operatorname{education}_{\operatorname{4-Year}}) - 0.73(\operatorname{education}_{\operatorname{Post-Grad}}) \\
\log\left[ \frac { P( \operatorname{approval}  \leq  \operatorname{Disapprove\ /\ Somewhat\ Disapprove} ) }{ 1 - P( \operatorname{approval}  \leq  \operatorname{Disapprove\ /\ Somewhat\ Disapprove} ) } \right] &= 2.73 + 0.61(\operatorname{ideology}_{\operatorname{Liberal}}) + 2.4(\operatorname{ideology}_{\operatorname{Moderate}}) + 4.46(\operatorname{ideology}_{\operatorname{Conservative}}) + 5.45(\operatorname{ideology}_{\operatorname{Very\ Conservative}}) - 0.1(\operatorname{education}_{\operatorname{High\ School\ Graduate}}) - 0.31(\operatorname{education}_{\operatorname{Some\ College}}) - 0.32(\operatorname{education}_{\operatorname{2-Year}}) - 0.52(\operatorname{education}_{\operatorname{4-Year}}) - 0.73(\operatorname{education}_{\operatorname{Post-Grad}}) \\
\log\left[ \frac { P( \operatorname{approval}  \leq  \operatorname{Neither\ Approve\ nor\ Disapprove} ) }{ 1 - P( \operatorname{approval}  \leq  \operatorname{Neither\ Approve\ nor\ Disapprove} ) } \right] &= 2.82 + 0.61(\operatorname{ideology}_{\operatorname{Liberal}}) + 2.4(\operatorname{ideology}_{\operatorname{Moderate}}) + 4.46(\operatorname{ideology}_{\operatorname{Conservative}}) + 5.45(\operatorname{ideology}_{\operatorname{Very\ Conservative}}) - 0.1(\operatorname{education}_{\operatorname{High\ School\ Graduate}}) - 0.31(\operatorname{education}_{\operatorname{Some\ College}}) - 0.32(\operatorname{education}_{\operatorname{2-Year}}) - 0.52(\operatorname{education}_{\operatorname{4-Year}}) - 0.73(\operatorname{education}_{\operatorname{Post-Grad}}) \\
\log\left[ \frac { P( \operatorname{approval}  \leq  \operatorname{Approve\ /\ Somewhat\ Approve} ) }{ 1 - P( \operatorname{approval}  \leq  \operatorname{Approve\ /\ Somewhat\ Approve} ) } \right] &= 3.93 + 0.61(\operatorname{ideology}_{\operatorname{Liberal}}) + 2.4(\operatorname{ideology}_{\operatorname{Moderate}}) + 4.46(\operatorname{ideology}_{\operatorname{Conservative}}) + 5.45(\operatorname{ideology}_{\operatorname{Very\ Conservative}}) - 0.1(\operatorname{education}_{\operatorname{High\ School\ Graduate}}) - 0.31(\operatorname{education}_{\operatorname{Some\ College}}) - 0.32(\operatorname{education}_{\operatorname{2-Year}}) - 0.52(\operatorname{education}_{\operatorname{4-Year}}) - 0.73(\operatorname{education}_{\operatorname{Post-Grad}})
\end{aligned}$$


```{r}
#| cache: true
```

```{r}
fit_approval |> tidy(conf.int = TRUE)
tidy_approval |> 
  select(term, estimate, conf.low, conf.high) |> 
  gt() |> 
  cols_label(
    term = "Variable",
    estimate = "Estimate",
    conf.low = "Lower CI",
    conf.high = "Upper CI"
  )
```
Public opinion about political leaders often depends on where people fall on the ideological spectrum. Using data from the 2020 Cooperative Election Study, one of the largest political surveys in the U.S., we explore how presidential approval varies across different ideological groups. One concern is that survey respondents might be systematically different from other Americans.